\documentclass[fleqn, 12pt]{article}
\setcounter{secnumdepth}{4}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{commath, esdiff}
\usepackage{datetime}
\usepackage{graphicx, epstopdf}
\usepackage{ulem}
\usepackage{xfrac}
\usepackage{enumerate}
\usepackage{tikz}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\theoremstyle{definition}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\newenvironment{solution}
{\begin{proof}[Solution]\let\qed\relax}
	{\end{proof}}

%opening
\title{Recitation 11}
\author{}
\date{\formatdate{7}{1}{2015}}

\begin{document}

\maketitle
%\setlength{\mathindent}{0pt}

\tableofcontents

\newpage
\section{Functions of Two Variables}

\subsection{Domain of Definition}

\begin{example}
	Find the domain of definition of 
	\begin{equation*}
		f(x,y) = \sqrt{(x^2 + y^2 - 16) (x^2 + y^2 - 9)}
	\end{equation*}
\end{example}

\begin{solution}
	\begin{align*}
		x^2 + y^2 - 16 &\geq 0 &\&&& x^2 + y^2 - 9 &\geq 0\\
		\intertext{or}
		x^2 + y^2 - 16 &\leq 0 &\&&& x^2 + y^2 - 9 &\leq 0
	\end{align*}
	Therefore,
	\begin{align*}
		\sqrt{x^2 + y^2} &\geq 4\\
		\intertext{and}
		\sqrt{x^2 + y^2} &\leq 3
	\end{align*}
\end{solution}

\subsection{Disproving Existence of Limits}

\begin{example}
	Does $\lim\limits_{(x,y) \to (0,0)} \dfrac{x^2 - y^2}{x^2 + y^2}$ exist?
\end{example}

\begin{solution}
	\begin{align*}
		\lim\limits_{(x,y) \stackrel{y = kx}{\to} (0,0)} \dfrac{x^2 - y^2}{x^2 + y^2} &= \lim\limits_{x \to 0} \dfrac{x^2 - k^2 x^2}{x^2 + k^2 x^2}\\
		&= \dfrac{1 - k^2}{1 + k^2}
	\end{align*}
	Therefore, the limit is different for different values of $k$. Hence, the limit does not exist.
\end{solution}

\begin{example}
	Does $\lim\limits_{(x,y) \to (0,0)} \dfrac{x y}{x^3 + y^{\sfrac{3}{2}}}$ exist?
\end{example}

\begin{solution}
	\begin{align*}
		\lim\limits_{(x,y) \stackrel{y = kx}{\to} (0,0)} \dfrac{x y}{x^3 + y^{\sfrac{3}{2}}} &= 0
	\end{align*}
	However, this is not enough to prove the existence of the limit.
	\begin{align*}
		\lim\limits_{(x,y) \stackrel{y = k x^2}{\to} (0,0)} \dfrac{x y}{x^3 + y^{\sfrac{3}{2}}} &= \dfrac{k}{1 + k^{\sfrac{3}{2}}}
	\end{align*}
	Therefore, the limit does not exist.
\end{solution}

\subsection{Proving Existence of Limits}

\begin{example}
	Check the continuity of 
	\begin{equation*}
		f(x,y) =
			\begin{cases}
				\dfrac{x^2 y}{x^2+ y^2} &;\quad (x,y) \neq (0,0)\\
				0 &;\quad (x,y) = (0,0)\\
			\end{cases}
	\end{equation*}
\end{example}

\begin{solution}
	\begin{equation*}
		0 \leq \left\lvert \dfrac{x^2 y}{x^2 + y^2} \right\rvert = \dfrac{x^2}{x^2 + y^2} |y| \leq |y|
	\end{equation*}
	Therefore, by the Sandwich Theorem,
	\begin{equation*}
		0 \leq \lim\limits_{\substack{
			x \to 0\\
			y \to 0}} \dfrac{x^2 y}{x^2 + y^2} |y| \leq \lim\limits_{y \to 0} |y|
	\end{equation*}
	Therefore,
	\begin{align*}
		\lim\limits_{(x,y) \to (0,0)} f(x) &= 0
	\end{align*}
	Therefore, $f(x)$ is continuous.
\end{solution}

\begin{example}
	Check for which values of $m$ the function
	\begin{equation*}
		f(x,y) =
			\begin{cases}
				\dfrac{x^m y}{x^2 + 4 y^2} &;\quad (x,y) \neq  (0,0)\\
				0 &;\quad (x,y) = (0,0)\\
			\end{cases}
	\end{equation*}
	is continuous at $(0,0)$.
\end{example}

\begin{solution}
	Let
	\begin{align*}
		x &= r \cos t\\
		y &= r \sin t
	\end{align*}
	Therefore, when $(x,y) \to (0,0)$, $r \to 0$.\\
	Therefore,
	\begin{align*}
		\lim\limits_{\substack{
			x \to 0\\
			y \to 0}} f(x,y) &= \lim\limits_{r \to 0} f(r \cos t, r \sin t)\\
		&= \lim\limits_{r \to 0} \dfrac{r^m \cos^m t \cdot r \sin t}{r^2 \cos^2 t + 4 r^2 \sin^2 t}\\
		&= \lim\limits_{r \to 0} \dfrac{r^{m+1} \cos^m t \sin t}{r^2 (\cos^2 t + 4 \sin^2 t)}\\
		&= \lim\limits_{r \to 0} \dfrac{r^{m-1} \cos^m t \sin t}{1 + 3 \sin^2 t}
	\end{align*}
	If the limit depends on $t$, the limit is path dependent. Hence, the limit does not exist.\\
	If $m > 1$,
	\begin{align*}
		\lim\limits_{(x,y) \to (0,0)} f(x,y) &= \lim\limits_{r \to 0} \dfrac{r^{m-1} \cos^m t \sin t}{1 + 3 \sin^2 t}\\
		&= 0
	\end{align*}
	Therefore, the limit is 0, and $f(x,y)$ is continuous at $(0,0)$.\\
	If $m = 1$,
	\begin{align*}
		\lim\limits_{(x,y) \to (0,0)} f(x,y) &= \lim\limits_{r \to 0} \dfrac{\cos t \sin t}{1 + 3 \sin^2 t}		
	\end{align*}
	The limit depends on $t$, and hence does not exist.\\
	If $m < 1$,
	\begin{align*}
		\lim\limits_{(x,y) \to (0,0)} f(x,y) &= \lim\limits_{r \to 0} \dfrac{r^{m-1} \cos^m t \sin t}{1 + 3 \sin^2 t}\\
		&= \infty
	\end{align*}
	Therefore, the limit does not exist.
\end{solution}

\end{document}